Ensembler: Enabling high-throughput molecular simulations at the superfamily scale

### 2015-04-15 manuscript/ms.tex
Fixed DFG-flip citation

Some amino acids can exist in different protonation states, depending on pH and on their local environment.
These protonation states can have important effects on biological processes.
For example, long timescale MD simulations have suggested that the conformation of the DFG motif of the TK Abl1---believed to be an important regulatory mechanism [CITE: Abl1 DFG flip evidence]---is controlled by protonation of the aspartate~\cite{shan:pnas:2009:abl}.
Currently, protonation states are assigned simply based on pH (a user-controllable parameter).
At neutral pH, histidines have two protonation states which are approximately equally likely, and in this situation the selection is therefore made based on which state results in a better hydrogen bond.
It would be highly desirable to instead use a method which assigns amino acid protonation states based on a rigorous assessment of the local environment.
We thus plan to implement an interface and command-line function for assigning protonation states with MCCE2~\cite{alexov-gunner:biophys-j:1997:mcce2,georgescu-alexov-gunner:biophys-j:2002:mcce2,song-mao-gunner:j-comput-chem:2009:mcce2}, which uses electrostatics calculations combined with Monte Carlo sampling of side chain conformers to calculate pKa values.

>>>>

Some amino acids can exist in different protonation states, depending on pH and on their local environment.
These protonation states can have important effects on biological processes.
For example, long timescale MD simulations have suggested that the conformation of the DFG motif of the TK Abl1---believed to be an important regulatory mechanism~\cite{nagar:cell:2003:abl-1opl}---is controlled by protonation of the aspartate~\cite{shan:pnas:2009:abl}.
Currently, protonation states are assigned simply based on pH (a user-controllable parameter).
At neutral pH, histidines have two protonation states which are approximately equally likely, and in this situation the selection is therefore made based on which state results in a better hydrogen bond.
It would be highly desirable to instead use a method which assigns amino acid protonation states based on a rigorous assessment of the local environment.
We thus plan to implement an interface and command-line function for assigning protonation states with MCCE2~\cite{alexov-gunner:biophys-j:1997:mcce2,georgescu-alexov-gunner:biophys-j:2002:mcce2,song-mao-gunner:j-comput-chem:2009:mcce2}, which uses electrostatics calculations combined with Monte Carlo sampling of side chain conformers to calculate pKa values.
